---
title: "CEO"
author: "Edward"
date: "26/10/2023"
output: html_document
---

```{r}
require(data.table)
require(matrixStats)
require(dplyr)
require(tidyr)
require(ggplot2)
require(annmatrix)
require(patchwork)
# source("code/environmentSetup.R")
source("code/fns.R")

# devtools::load_all("rsrc")
sincos2acro <- function(sin,cos, per=24) (atan2(sin,cos)/2/pi*per+per)%%per
sincos2amp <- function(sin,cos) 2*sqrt(sin^2+cos^2)
```

### Import data

```{r fig.height = 3,fig.width =4}
mydat <- readRDS("output/methylation_stats.RDS")
dat <- rowanns(mydat) 
# source("code/fns.R")

```

### Figure 1

#### (a) (b)
```{r fig.height =2,fig.width =6}
qPCR <- fread("input/170217_qPCR_logddCT.csv")
qPCR <- qPCR %>% rename(Gene = Target)
qPCR[qPCR$Gene == "PER1",]$Gene <- "Per1"
qPCR[qPCR$Gene == "ARNTL",]$Gene <- "Arntl"
qPCR[qPCR$Gene == "PER2",]$Gene <- "Per2"

## Fixing CT: continuous time 
qPCR <- qPCR %>% rename(ct = ZT, time_of_day=CT) %>%
  mutate(zt = ct2zt(ct))

facs <- fread("input/flow_analysis.csv")
facs$zt <- ct2zt(facs$ct)

panel_a <- qPCR %>% 
  subset(Gene != "VIL1" & Gene != "LCT") %>% 
  ggplot(aes(zt, logddCt, color=Gene))+
  theme_bw() +
  annotate("rect",xmin = 12,  xmax = Inf,  ymin = -Inf,  ymax = Inf, fill = "grey88", col="white", alpha = 0.4) +
  geom_point(alpha=0.5, size=0.5)+
  geom_smooth(method="lm", se=F, alpha=0.3,formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi)) +
  scale_x_continuous(breaks=seq(0,21,3)) + ylab("") + 
  xlab("ZT") + 
  ylab("relative expression") +
  scale_color_manual(values=c(colkey[5],colkey[10], colkey[15]))   +
  facet_wrap(~Gene, scales="free", nrow=1) +
  ggtitle("a") + 
  theme(
    strip.background =element_blank(),
                # legend.title = element_blank(),
    legend.position = "none",
    #         legend.position = c(0.5,1.2),
    # legend.key.size = unit(0.05,'cm'),
    # legend.direction = "horizontal",
        text = element_text(size = 7))

panel_b <-
  facs %>%
  subset(Var1 == "G2/M" | Var1 == "S") %>%
  ggplot(aes(zt, perCount, color=Var1))+
  annotate("rect",xmin = 12,  xmax = Inf,  ymin = -Inf,  ymax = Inf, fill = "grey88", col="white", alpha = 0.4) +
  geom_point(alpha=0.5, size=0.5)+
  geom_smooth(method="lm", se=F, alpha=0.3,formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi)) + 
  scale_x_continuous(breaks=seq(0,21,3)) +
  scale_color_manual(values=rev(c("grey25", "darkred")), name="Cell cycle\nstage") + ylab("cell proportion (%)")+
  # scale_color_grey() +
  xlab("ZT") +
  guides(fill=guide_legend(title="Cell cycle stage"))  + 
  ggtitle("b") + 
  facet_wrap(~Var1, scales="free", nrow=1) +
  theme(strip.background =element_blank(),
        legend.title = element_blank(),
                legend.position = "none",
    #         legend.position = c(0.5,1.2),
    # legend.key.size = unit(0.05,'cm'),
    # legend.direction = "horizontal", 
        text = element_text(size = 7))
  
```

<!-- #### (c) -->
<!-- ```{r fig.height =2,fig.width =3} -->

<!-- panel_c <- facs %>% -->
<!--   mutate(Var1 = ifelse(grepl("G1", Var1), "G0/G1", Var1)) %>% -->
<!--   ggplot(aes(Var1, perCount)) + -->
<!--   geom_boxplot() + ylab("cell proportion (%)")+ -->
<!--   xlab("cell cycle phase")  + -->
<!--   ggtitle("c") + -->
<!--   theme(strip.background =element_blank(), -->
<!--         text = element_text(size = 7)) -->

<!-- facs %>% -->
<!--   filter(Var1 == "S") %>% -->
<!--   ggplot(aes(x=zt, y=perCount)) + -->
<!--   geom_hline(yintercept = 15.46999, col="Red") + -->
<!--   geom_hline(yintercept = 16.25178) + -->
<!--   geom_point() -->

<!-- median(facs[facs$Var1 == "S",]$perCount) -->

<!-- quantile(facs[facs$Var1 == "S",]$perCount, 0:5/5) -->
<!-- sum(facs[facs$Var1 == "S",]$perCount)/2 -->
<!-- facs[facs$Var1 == "S",] %>% -->
<!--   filter(between(zt, 10,24)) %>% -->
<!--   summarise(sum(perCount)) -->
<!-- unique(facs$zt) -->

<!-- facs %>% -->
<!--   filter(Var1 == "S") %>% -->
<!--   mutate(major = between(zt, 13, 19)) %>% -->
<!--   group_by(major) %>% -->
<!--   summarise(med = median(perCount), sum=sum2(perCount), sum(freq)) -->

<!-- ``` -->

#### (c) osc heatmap
Heatmap of osc-modC densities (range=0.9-1.10; mean-centered). 

```{r fig.height =5,fig.width = 2.5}

HReff <- annapply(mydat, c("media"), function(x) {
		fit <- annOLSwFtest(x, ~sin+cos + as.factor(age)*ct, ~as.factor(age)*ct,ver=T)
		data.frame("value"=fit$pvalue,
		           fit$fit$betas,
		           rowanns(x))
	}) %>%
  separate(grp, c("media"), remove=FALSE) %>% 
  mutate(acro=sincos2acro(sin,cos)) %>%
  mutate(amp=sincos2amp(sin,cos))

osc_order <- HReff %>%
  filter(grp == "ENR") %>%
  filter(value < 0.05) %>%
  mutate(slp = -log10(value)*amp) %>%
  # arrange(ct2zt_e3e4(acro)) %>%
  arrange(acro)

plotdf <- data.frame(modid = mydat@modid, mydat)

plotdfx <- plotdf %>%
  filter(modid %in% osc_order$modid) %>%
  melt(., id.var="modid") %>%
  rename(sampleid = variable) %>%
  left_join(., data.frame(zt = mydat$zt, ct=mydat$ct, age= mydat$age, media=mydat$media, sampleid= paste0("X", mydat$sampleid)), by = "sampleid") %>%
  filter(media !="ENRIV") %>%
  group_by(modid, age) %>%
  mutate(mean_value = mean(value)) %>%
  # filter(!is.na(mean_value)) %>%
  mutate(centered = value - mean_value) %>%
  ungroup() %>%
  group_by(modid, ct) %>%
  mutate(key = 1:n()) %>%
  mutate(key = paste(zt, key, sep=".")) %>%
  ungroup() %>%
  mutate(modid = factor(modid, c(osc_order$modid))) %>%
  mutate(key = factor(key, c(sort(as.numeric(as.character(unique(key))))))) %>%
  mutate(grp = "bisulfite-seq\n(osc-modCs)")

scale_bin <- quantile(plotdfx$centered*100, c(0.05, 0.95), na.rm=TRUE)

panel_c <- plotdfx %>%
  ggplot(aes(x=(key), y=modid, fill=(centered)*100)) +
  geom_tile() + theme(axis.text.x = element_text(angle = 90, hjust = 1),
                      axis.text.y = element_blank()) +
  viridis::scale_fill_viridis(option="mako", limits=c(scale_bin[1], scale_bin[2]), oob=scales::squish) +
  # scale_fill_continuous(colours = terrain.colors()) +
  scale_x_discrete(labels=c( 0, "", "",4, "","", 8, "","", 12, "","", 16, "", "",20, "")) +
  xlab("zeitgeber time") + ylab("") + scale_y_discrete(breaks=NULL) +
  labs(fill='normalized\n%modC') +
  facet_grid(~grp)  + 
  ggtitle("c") + 
  theme(strip.background =element_blank(),
        text = element_text(size = 7))

```

#### (d) osc p-value histogram
```{r fig.height = 2,fig.width =3}

HReff <- annapply(mydat, c("media"), function(x) {
		fit <- annOLSwFtest(x, ~sin+cos + as.factor(age)*ct, ~as.factor(age)*ct,ver=T)
		data.frame("value"=fit$pvalue,
		           fit$fit$betas,
		           rowanns(x))
	}) %>%
  separate(grp, c("media"), remove=FALSE) %>% 
  mutate(acro=sincos2acro(sin,cos)) %>%
  mutate(amp=sincos2amp(sin,cos))

panel_d <-
  HReff %>%
  # filter(evc_pval < 0.05) %>%
  filter(grp == "ENR") %>%
  ggplot(aes(x=value)) +
  geom_histogram(breaks=0:20/20, col="white", size=0.2) +
  xlab("oscillation p-value")  + 
  ggtitle("d") + 
  # coord_cartesian(ylim = c(1000,2000)) +
  # ylim(0,2000) +
  theme(strip.background =element_blank(),
        text = element_text(size = 7))

```

<!-- ### (runs test) -->
<!-- ```{r} -->
<!-- mydat[,mydat$media == "ENR"] -->
<!-- fit <- annOLSwFtest(mydat[,mydat$media == "ENR"], ~sin+cos + as.factor(age)*ct, ~as.factor(age)*ct,ver=T) -->

<!-- # randtests::runs.test(fit$fit) -->
<!-- # fit$fit$fit$residuals -->

<!-- mat_runs <- function(mat){ -->

<!--   # same for all rows -->
<!--   nm=ncol(mat) -->
<!--   # stopifnot(nm >= 30)  -->
<!--   warning("To Matt: Add back the stopping criteria at some point!") -->
<!--   denom <- (nm^2 * (nm - 1)) -->

<!--   # Convert to runs -->
<!--   mat <- ((mat > 0)*1) -->
<!--   runs = rowSums(mat[,-1] != mat[,-ncol(mat)])+1 -->
<!--   ms=rowSums(mat==0) -->
<!--   ns=rowSums(mat==1) -->

<!--   E <- 1 + 2*ns*ms / nm -->
<!--   s2 <- (2*ns*ms * (2*ns*ms - ns - ms)) / denom -->
<!--   statistic <- (runs - E) / sqrt(s2) -->
<!--   p.value = pnorm(statistic) -->
<!--   return(p.value) -->
<!-- } -->

<!-- unique(mydat$age) -->
<!-- tmp_dat <- mydat[,mydat$media == "ENR" & mydat$age == "62"] -->
<!-- # fit <- annOLSwFtest(mydat[,mydat$media == "ENR"], ~sin+cos + as.factor(age)*ct, ~as.factor(age)*ct,ver=T) -->
<!-- fit <- annOLSwFtest(tmp_dat, ~sin+cos + ct, ~ ct, ver=T) -->
<!-- tmp_df <- fit$fit$residuals -->

<!-- neworder <- tmp_df$'' %>% -->
<!--   arrange(ct) -->

<!-- tmp_df2 <- tmp_df[,match(neworder$sampleid, tmp_df$sampleid)] -->
<!-- colnames(tmp_df2) == neworder$sampleid -->
<!-- tmp_df2$sampleid == neworder$sampleid -->
<!-- tmp_df2$ct -->

<!-- runs_p <- mat_runs(tmp_df2) -->
<!-- hist(runs_p) -->

<!-- data.frame(value=fit$pvalue) %>% -->
<!--   # filter(media == "ENR") %>% -->
<!--   mutate(runsP = runs_p) %>% -->
<!--   ggplot(aes(x=value, fill=runsP<0.05)) + -->
<!--   geom_histogram(breaks=0:20/20, col="white", size=0.2) + -->
<!--   xlab("harmonic regression p-value")  +  -->
<!--   ggtitle("e") +  -->
<!--   theme(strip.background =element_blank(), -->
<!--         text = element_text(size = 7)) -->
<!-- ``` -->


#### (e) osc permutations
```{r}

plotPerms <- function(realdf,permdf){

    realres <- realdf %>% group_by(grp) %>% summarise(pct=mean(pvalue<0.05,na.rm=T))
    realres$iter="real"
    permres <- permdf %>% group_by(grp,iter) %>% summarise(pct=mean(pvalue<0.05,na.rm=T)) %>% ungroup()
    pvals=rbind(permres,realres) %>% group_by(grp) %>% summarise(p=sum(pct[iter=="real"]<=pct[iter!="real"])/sum(iter!="real"))
    master<- merge(realres,pvals)
    
    permres %>% 
        ggplot(aes(grp,pct))+
        geom_point(position=position_jitter(width=0.2),color="grey")+
        coord_flip()+
        geom_hline(yintercept = 0.05,linetype=2)+
        geom_point(data=master,color="red",size=3)+
        geom_text(data=master,aes(grp,pct,label=p))+
        ylab("%Significant (p<0.05)")+xlab("")
}

annOLSwFtest_permSinCos <- function(x,f1,f2,N){
    lapply(1:N,function(i){
        set.seed(i)
        tmpdat <- x
        colanns(tmpdat) <- colanns(tmpdat) %>% 
            mutate(shuf=sample(1:length(sin))) %>% 
            mutate(sin=sin[shuf],cos=cos[shuf])
        data.frame(pvalue=annOLSwFtest(tmpdat,f1,f2))
    }) %>% data.table::rbindlist(idcol="iter")
}

permANOVAwithin_SinCos <- function(dat,f1,f2,N,shufvar="sampleid"){
    permdf <- annapply(dat,c("age","media"),function(x){
        annOLSwFtest_permSinCos(x,f1,f2,N)
    })
    
    realdf <- annapply(dat,c("age","media"),function(x){
        res <- annOLSwFtest(x,f1,f2)
        data.frame(pvalue=res,iter="real")
    })
  
    p <- plotPerms(realdf,permdf)
    
    print(p)
    
    return(list("perms"=permdf,"real"=realdf,p))
}

permANOVAwithin_SinCos_media <- function(dat,f1,f2,N,shufvar="sampleid"){
    permdf <- annapply(dat,c("media"),function(x){
        annOLSwFtest_permSinCos(x,f1,f2,N)
    })
    
    realdf <- annapply(dat,c("media"),function(x){
        res <- annOLSwFtest(x,f1,f2)
        data.frame(pvalue=res,iter="real")
    })
  
    p <- plotPerms(realdf,permdf)
    
    print(p)
    
    return(list("perms"=permdf,"real"=realdf,p))
}

# aovres <- readRDS("output/ReliableCytosines.RDS")
# mydatA <- mydat[aovres$pvalue<0.05,]
# mydat@aov_pval <- aovres$pvalue

Nperm=1000
# Run permutations
cachefile <- paste0("output/cache/HRpermplotDat_media_all",Nperm,".RDS")
if(file.exists(cachefile)){
    HRpermplotDat <- readRDS(cachefile)
    HRpermplotDat[[3]]
}else{
    HReff_perms <- permANOVAwithin_SinCos_media(mydat,~sin+cos+as.factor(age)*ct,~as.factor(age)*ct,Nperm)
    
    # Combine permutations by media
    realres <- HReff_perms$real %>% 
        mutate(age=substr(grp,1,7)) %>% 
        mutate(media=substr(grp,9,20)) %>%
        mutate(iter="real") %>% 
        group_by(media,iter) %>% 
        summarise(pct=mean(pvalue<0.05,na.rm=T)) %>% 
        ungroup()
    
    permres <- HReff_perms$perms %>% 
        mutate(age=substr(grp,1,7)) %>% 
        mutate(media=substr(grp,9,20)) %>%
        mutate(iter=as.character(iter)) %>% 
        group_by(media,iter) %>% 
        summarise(pct=mean(pvalue<0.05,na.rm=T)) %>% 
        ungroup()
    
    pvals=rbind(permres,realres) %>% group_by(media) %>% summarise(p=sum(pct[iter=="real"]<=pct[iter!="real"])/sum(iter!="real"))

    master <- merge(pvals,realres)
    
    HRpermplotDat <- list("pctsig_perms"=permres,
                                                "pvals"=master,
                                                "plot"=HReff_perms[[3]],
                                                HReff_perms$real)
    
    saveRDS(HRpermplotDat,file=cachefile,compress=F)
}

HRpermplotDat[[1]] %>% 
    ggplot(aes(media,pct))+
    geom_point(position=position_jitter(width=0.2),color="grey")+
    # facet_wrap(~media,scales="free")+
    coord_flip()+
    geom_point(data=HRpermplotDat[[2]],color="red",size=3)+
    geom_text(data=HRpermplotDat[[2]],aes(media,pct,label=p))

panel_e <- HRpermplotDat[[1]] %>% 
    ggplot(aes(pct))+
  geom_histogram( col="white", size=0.2) +
    geom_vline(xintercept = HRpermplotDat[[2]]$pct, col="darkred") +
    ylab("permutations") + xlab("mean r-squared") +
   ggtitle("e") + 
    # facet_wrap(~media,scales="free")+
    # coord_flip()+
    # geom_point(data=HRpermplotDat[[2]],color="red",size=3)+
    geom_text(data=HRpermplotDat[[2]],size=2, aes(x=HRpermplotDat[[2]]$pct+0.005, y=100,pct,label=paste0("p=",p)))  + 
  theme(strip.background =element_blank(),
        text = element_text(size = 7))
```

#### (f) osc acrophase
```{r fig.height = 2,fig.width =3}

### Convert ct acrophases to zt to be more applicable to biology

panel_f <- HReff %>%
  filter(grp == "ENR") %>%
  filter(value < 0.05) %>%
  # ggplot(aes(x=ct2zt_e3e4(acro))) +
  ggplot(aes(x=acro)) +
  geom_rect(aes(xmin=12,xmax=Inf,ymin=0,ymax=Inf), alpha=0.1, fill="lightgrey", col="white") +
  geom_histogram(breaks=0:24, col="white", size=0.2) +
  scale_x_continuous(breaks=seq(0,24,4)) +
  # coord_cartesian(ylim = c(1000,2000)) +
  xlab("acrophase")  + 
  ggtitle("f") + 
  theme(strip.background =element_blank(),
        text = element_text(size = 7))

```

#### (g) pseudo Chisqr test 
```{r fig.height = 2,fig.width =9}
chi_tmp <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & MPseudo2_ENR_pval < 0.05) %>%
  mutate(acro_bins = cut(MHR_ENR_xCOLL_acro, seq(0,24,3))) %>%
  mutate(direction = ifelse(MPseudo2_ENR_est < 0, "loss","gain")) %>%
  select(acro_bins, direction)

chi_res <- chisq.test(table(chi_tmp$acro_bins, chi_tmp$direction))
graphed <- chi_grapher(chi_res, two_by_two = FALSE)
p1 <- graphed[[4]] + ggtitle(paste("interaction p=",signif(chi_res$p.value, 2), "\n\nnumber of osc-modCs", sep=""))
p3 <- graphed[[3]] + ggtitle(" \n\n% contribution") 
# gridExtra::grid.arrange(p1, p3, nrow=1)

```

```{r fig.height = 2.25,fig.width =3.5}

compounded <- data.frame(100*chi_res$residuals^2/chi_res$statistic) %>%
      mutate(Freq = Freq*sign(data.frame(chi_res$residuals)$Freq)) 
    
panel_g <- compounded %>%
  filter(Freq > 0) %>%
  mutate(Freq = ifelse(Var2 == "loss", -Freq, Freq)) %>%
  mutate(Var1 = factor(Var1, levels=rev(sort(unique(Var1))))) %>%
  # mutate(cols = paste(Var2, Var1)) %>%
  ggplot(aes(x=Var1, y=Freq, fill=Var1)) +
  geom_bar(stat="identity") + coord_flip() +
  # scale_fill_viridis_d() +
  scale_fill_grey() +
theme(legend.position = "none") +
  xlab("oscillation\nacrophases") +
  ylab("% contribution to linear trend") +
  scale_y_continuous(breaks=seq(-10, 15, 5), labels=abs(seq(-10, 15, 5)))   + 
  ggtitle("g") + 
  theme(strip.background =element_blank(),
        text = element_text(size = 7))

```



#### (h) ex vivo Chisqr test 
```{r fig.height = 2,fig.width =9}
chi_tmp <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval < 0.05) %>%
  mutate(acro_bins = cut(MHR_ENR_xCOLL_acro, seq(0,24,3))) %>%
  mutate(direction = ifelse(EXvivo_ENR_xCOLL_est < 0, "loss","gain")) %>%
  select(acro_bins, direction)

chi_res <- chisq.test(table(chi_tmp$acro_bins, chi_tmp$direction))
graphed <- chi_grapher(chi_res, two_by_two = FALSE)
p1 <- graphed[[4]] + ggtitle(paste("interaction p=",signif(chi_res$p.value, 2), "\n\nnumber of osc-modCs", sep=""))
p3 <- graphed[[3]] + ggtitle(" \n\n% contribution") 
# gridExtra::grid.arrange(p1, p3, nrow=1)

```

```{r fig.height = 2.25,fig.width =3.5}

compounded <- data.frame(100*chi_res$residuals^2/chi_res$statistic) %>%
      mutate(Freq = Freq*sign(data.frame(chi_res$residuals)$Freq)) 
    
panel_h <- compounded %>%
  filter(Freq > 0) %>%
  mutate(Freq = ifelse(Var2 == "loss", -Freq, Freq)) %>%
  mutate(Var1 = factor(Var1, levels=rev(sort(unique(Var1))))) %>%
  # mutate(cols = paste(Var2, Var1)) %>%
  ggplot(aes(x=Var1, y=Freq, fill=Var1)) +
  # annotate("rect", xmin = "(12,15]", xmax="(21,24]", ymin=-Inf, ymax=Inf ) +
  geom_bar(stat="identity") + coord_flip() +
  # scale_fill_viridis_d() +
  scale_fill_grey() +
  theme(legend.position = "none") +
  xlab("oscillations\nacrophases") +
  ylab("% contribution to linear trend") +
  scale_y_continuous(breaks=seq(-20, 5, 5), labels=abs(seq(-20, 5, 5)))  + 
  ggtitle("h") + 
  theme(strip.background =element_blank(),
        text = element_text(size = 7))


```

#### meta (g)
```{r fig.height = 4,fig.width =4}
panel_meta_g <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & MPseudo2_ENR_pval < 0.05) %>%
  ggplot(aes(x=MHR_ENR_xCOLL_acro, y=MPseudo2_ENR_est)) +
  geom_density_2d(col="darkgrey", alpha=0.5) +
  # ylim(-0.0025, 0.0025) +
  ylab("donor age\n(%modC/day)") + 
  xlab("acrophase") +
  ggtitle("g") +
  # ylim(0.1,-0.1) +
  geom_hline(yintercept = 0, col="grey") +
  scale_y_continuous(labels = scales::percent, limits = c(-0.0005, 0.0005)) +
  scale_x_continuous(breaks=seq(0,24,4)) +
  # geom_point(size=0.5, alpha=0.1) +
  annotate("text", size=2, x=12, y=0.0020, label=paste0("R2=0.02; p=0.002; n=603"))  + 
  geom_smooth(method="lm", se=F, size=1, alpha=1, col="black",
              formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi), fullrange=TRUE) + 
  theme(strip.background =element_blank(),
        text = element_text(size = 7))

pseudo_dat <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & MPseudo2_ENR_pval < 0.05)

res <- lm(MPseudo2_ENR_est ~ sin(MHR_ENR_xCOLL_acro/24*2*pi) + cos(MHR_ENR_xCOLL_acro/24*2*pi), data=pseudo_dat)
summary(res)
```

#### meta (h)
```{r fig.height = 4,fig.width =4}
panel_meta_h <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval < 0.05) %>%
  ggplot(aes(x=MHR_ENR_xCOLL_acro, y=EXvivo_ENR_xCOLL_est)) +
  geom_density_2d(col="darkgrey", alpha=0.5) +
  # ylim(-0.0025, 0.0025) +
  ylab("Ex vivo aging\n(%modC/hour)") + 
  xlab("acrophase") +
  ggtitle("h") +
  geom_hline(yintercept = 0, col="grey") +
  scale_y_continuous(labels = scales::percent, limits = c(-0.001, 0.001)) +
  scale_x_continuous(breaks=seq(0,24,4)) +
  # geom_point(size=0.5) +
  annotate("text", size=2, x=12, y=0.0020, label=paste0("R2=0.1; p=3.1e-06; n=270"))  + 
  geom_smooth(method="lm", se=F, size=1, alpha=1, col="black",
              formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi), fullrange=TRUE) + 
  theme(strip.background =element_blank(),
        text = element_text(size = 7))

exvivo_dat <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval < 0.05)

res <- lm(EXvivo_ENR_xCOLL_est ~ sin(MHR_ENR_xCOLL_acro/24*2*pi) + cos(MHR_ENR_xCOLL_acro/24*2*pi), data=exvivo_dat)
summary(res)
```

#### (i) osc-modCs: Significant only 
```{r fig.height = 2.5,fig.width =5}

oscmodc <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05) %>%
  # filter(MHR_ENR_xCOLL_pval < 0.05 & HR_tl_s4UCORCT_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  # mutate(acro = ifelse(MHR_ENR_xCOLL_acro < 12, "osc-modC\nacrophase\n(day)",
  #                      "osc-modC\nacrophase\n(night)")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  select(modid, zt, ct, pctm, age, media, acro) %>%
  filter(media == "ENR") %>%
  group_by(modid, media, age) %>%
  mutate(pctm_age = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm - pctm_age) 

random_cpgs <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05) %>%
    mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  group_by(acro) %>%
  sample_n(., 10)

panel_i <-
  oscmodc %>%
  # mutate(zt = zt + 24) %>%
  # rbind(., oscmodc) %>%
  # mutate(zt = zt + 24) %>%
  # rbind(., oscmodc) %>%
  filter(between(zt, 0, 25)) %>%
  filter(zt <= 24) %>%
  filter(modid %in% random_cpgs$modid) %>%
  # filter(acro == "(3,6]" | acro == "(6,9]" | acro == "(9,12]"| acro == "(18,21]" |acro == "(21,24]") %>%
  mutate(acro = factor(acro, levels=c("light", "dark", "NA"))) %>%
  # filter(acro == "acrophase:\nZT:0-12") %>%
  filter(!is.na(acro)) %>%
  ggplot(aes(x=zt, y=pctm_norm*100)) +
  annotate("rect",ymin=-Inf, ymax=Inf, xmin = 13, xmax=19, fill="grey", alpha=0.5) +
  # facet_wrap(acro~., nrow=1, strip.position = "top")  +
  theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  # geom_vline(xintercept=c(S_acro, S_acro+24), linetype="dotted", col="black") +
  geom_smooth(aes(group=as.factor(modid), col=acro),
              # col="#999999", 
              # size=0.2,
              method="lm", se=F, size=0.25, alpha=0.25,
              formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi), fullrange=TRUE) +
  scale_x_continuous(breaks=seq(0,24,4)) +
  # annotate("rect",ymin=-Inf, ymax=Inf,
  #          xmin = G2M_acro-G2M_acro_sd-24, xmax=G2M_acro+G2M_acro_sd-24, fill="#980043", alpha=0.1) +
  # annotate("rect",ymin=-Inf, ymax=Inf, 
  #          xmin = G2M_acro-G2M_acro_sd, xmax=G2M_acro+G2M_acro_sd, fill="#980043", alpha=0.1) +
  # annotate("rect",ymin=-Inf, ymax=Inf, 
  #          xmin = G2M_acro-G2M_acro_sd+24, xmax=G2M_acro+G2M_acro_sd+24, fill="#980043", alpha=0.1) +
  # geom_vline(xintercept=c(G2M_acro, G2M_acro+24, G2M_acro-24), linetype="dashed", col="#980043") +
  # annotate("rect",ymin=-Inf, ymax=Inf, xmin = 13+24, xmax=19+24, fill="grey", alpha=0.5) +
  # annotate("rect",ymin=-Inf, ymax=Inf, xmin = 10+48, xmax=22+48, fill="blue", alpha=0.2) +
  # annotate("rect",ymin=-Inf, ymax=Inf, xmin = S_acro-S_acro_sd+24, xmax=S_acro+S_acro_sd+24, fill="blue", alpha=0.2) +
  geom_smooth(aes(col=acro), method="lm", se=F, size=2, alpha=0.1, 
              formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi), fullrange=TRUE) +
  geom_vline(xintercept=c(S_acro), linetype="dotted", col="black") +
  ylab("normalized % modC") +
  xlab("ZT") +
  # ylim(-2.5,2.5) +
  # scale_color_manual(values=c(viridis::viridis(1000))) +
  # scale_color_grey() +
  scale_color_manual(values=c("darkred","darkblue")) +
  # scale_x_continuous(breaks=seq(0, 24, 4), labels=c(rep(seq(0, 20, 4),1),0))  + 
  ggtitle("i") + 
  theme(strip.background =element_blank(),
        text = element_text(size = 7))
```

##### MESOR calc
```{r}
oscmodc <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05) %>%
  # filter(MHR_ENR_xCOLL_pval < 0.05 & HR_tl_s4UCORCT_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  # mutate(acro = ifelse(MHR_ENR_xCOLL_acro < 12, "osc-modC\nacrophase\n(day)",
  #                      "osc-modC\nacrophase\n(night)")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  select(modid, zt, ct, pctm, age, media, acro) %>%
  filter(media == "ENR") %>%
  group_by(modid, media, age) %>%
  mutate(pctm_age = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm - pctm_age) 

MESORs <- oscmodc %>%
  ungroup() %>%
  group_by(acro) %>%
  summarise(mean = mean(pctm)*100,
            sd = sd(pctm)*100,
            n=n()) 

light <- oscmodc %>%
  filter(acro=="light") %>%
  mutate(zt= as.numeric(zt))

m1 <- cosinor::cosinor.lm(pctm ~ time(zt), data= light, period =24)

m1$coefficients

dark <- oscmodc %>%
  filter(acro=="dark") %>%
  mutate(zt= as.numeric(zt))
m2 <- cosinor::cosinor.lm(pctm ~ time(zt), data= dark, period =24)

m2$coefficients
```

<!-- #### (i) tmp -->
<!-- ```{r fig.height = 2.5,fig.width =5} -->

<!-- oscmodc <- dat %>% -->
<!--   filter(MHR_ENR_xCOLL_pval < 0.05) -->

<!-- oscmodc %>% -->
<!--   ggplot(aes(x=MHR_ENR_xCOLL_acro, y=EXvivo_ENR_xCOLL_est)) + -->
<!--   geom_point() +  -->
<!--   geom_smooth(method="lm", se=F, alpha=0.3,formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi)) + -->
<!--   geom_vline(xintercept=c(G2M_acro), linetype="dashed", col="#980043") + -->
<!--   geom_vline(xintercept=c(S_acro), linetype="dashed", col="blue") + -->
<!--   annotate("rect",ymin=-Inf, ymax=Inf, xmin = S_acro-S_acro_sd, xmax=S_acro+S_acro_sd, fill="blue", alpha=0.2) + -->
<!--   annotate("rect",ymin=-Inf, ymax=Inf, xmin = G2M_acro-G2M_acro_sd, xmax=G2M_acro+G2M_acro_sd, fill="#980043", alpha=0.2)  -->

<!-- ``` -->

#### (j) pctm at DNA synthesis
```{r fig.height = 3.5,fig.width =8}

oscmodc <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  # filter(EXvivo_ENR_xCOLL_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  mutate(direction = ifelse(EXvivo_ENR_xCOLL_est < 0, "linear loss", "linear gain")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark",
                          MHR_ENR_xCOLL_pval > 0.05 ~ "not oscillating")) %>%
  filter(media == "ENR") %>%
  select(modid, zt, ct, pctm, age, acro, direction) %>%
  group_by(modid, age, acro, direction) %>%
  mutate(center = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm-center) %>%
  filter(zt == 16) %>%
  mutate(zt = ifelse(zt == 16, "peak S phase (~ZT16)", "peak G2/M (~ZT23)")) %>%
  mutate(cat = paste0(acro," : ",direction)) %>%
  filter(cat == "dark : linear gain" | cat == "light : linear loss")

mean_lines <- oscmodc %>%
  ungroup() %>%
  group_by(cat) %>%
  summarise(mean = mean(pctm_norm)*100,
            sd = sd(pctm_norm)*100,
            n=n())

panel_j <-
oscmodc %>%
  ggplot(aes(x=(pctm_norm)*100, col=cat)) +
  geom_vline(xintercept = mean_lines$mean[1], col=rev(viridis::viridis(3))[2:3][1], alpha=1,linetype="dotted") +
  geom_vline(xintercept = mean_lines$mean[2], col=rev(viridis::viridis(3))[2:3][2], alpha=1,linetype="dotted") +
  geom_density(aes(fill=cat), alpha=0.5) +
  # scale_color_manual(values=rev(viridis::viridis(3))[2:3], name="") +
  # scale_fill_manual(values=rev(viridis::viridis(3))[2:3], name="") +
  # scale_fill_grey() +
  # scale_color_grey() +
  scale_color_manual(values=c("darkblue","darkred")) +
  scale_fill_manual(values=c("darkblue","darkred")) +
  # facet_wrap(~subheading) +
  xlab("normalized %modC") +
  ggtitle("j") + 
  theme(
   legend.position = c(0.5,1.18),
        legend.key.size = unit(0.05,'cm'),
        # legend.direction = "horizontal",
      legend.title = element_blank(),
        strip.background =element_blank(),
    legend.background = element_blank(),
        text = element_text(size = 7))

```


#### (k) Trying something new
```{r fig.height = 3.5,fig.width =8}
test <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  # filter(EXvivo_ENR_xCOLL_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  mutate(direction = ifelse(EXvivo_ENR_xCOLL_est < 0, "linear loss", "linear gain")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark",
                          MHR_ENR_xCOLL_pval > 0.05 ~ "not oscillating")) %>%
  filter(media == "ENR") %>%
  select(modid, zt, ct, EXvivo_ENR_xCOLL_est, pctm, age, acro, direction) %>%
  group_by(modid, age, acro, direction) %>%
  mutate(center = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm-center) %>%
  filter(zt == 16) %>%
  mutate(zt = ifelse(zt == 16, "peak S phase (~ZT16)", "peak G2/M (~ZT23)")) %>%
  mutate(cat = paste0(acro," : ",direction)) %>%
  filter(cat == "dark : linear gain" | cat == "light : linear loss")

panel_k <-
test %>%
  # filter(MHR_ENR_xCOLL_pval < 0.05) %>%
  # filter(EXvivo_ENR_xCOLL_est < 0) %>%
  # mutate(acro = cut(abs(MHR_ENR_xCOLL_acro), breaks=seq(0,24, 3))) %>%
  # mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 9~"ZT:0~9",
  #                         MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 21~"ZT:12~21")) %>%
  # mutate(acro = ifelse(is.na(acro),"other ZTs", acro)) %>%
  # mutate(slope_G2M = getSlope(S_acro, MHR_ENR_xCOLL_acro, MHR_ENR_xCOLL_amp)) %>%
  ggplot(aes(x=abs(EXvivo_ENR_xCOLL_est)*100, y=(pctm_norm*100), col=cat)) +
  geom_smooth(method="lm", size=0.5, fullrange=TRUE, se=F, formula='y ~ x') +
  geom_point(size=0.2, alpha=0.2) +
  # scale_color_manual(values=rev(viridis::viridis(3))[2:3], name="") +
  scale_color_manual(values=c("darkblue","darkred")) +
  scale_fill_manual(values=c("darkblue","darkred")) +
  # scale_fill_manual(values=rev(viridis::viridis(3))[2:3], name="") +
  ylab("deviation from MESOR\nat S phase (%modC)") +
  xlab("rate of linear trend\n(%modC/hour)")  + 
  # scale_x_log10() +
  # scale_y_log10() +
  # xlim(0,0.25) +
  ggtitle("k") +
  theme(
    legend.position = c(0.5,1.18),
    legend.key.size = unit(0.05,'cm'),
    legend.key = element_rect(colour = "transparent", fill = "transparent"),
    legend.title = element_blank(),
    legend.direction = "vertical", 
    strip.background =element_blank(),
        text = element_text(size = 7))

cor.test(test[test$cat == "light : linear loss",]$EXvivo_ENR_xCOLL_est, test[test$cat == "light : linear loss",]$pctm_norm)

cor.test(test[test$cat == "dark : linear gain",]$EXvivo_ENR_xCOLL_est, test[test$cat == "dark : linear gain",]$pctm_norm)

```

##### lm pvalue
```{r}
lm_res <- lm(oscmodc$pctm_norm ~ oscmodc$cat)
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}

lmp(lm_res)

```



### Figure collate
```{r fig.height=6, fig.width=6}

# (panel_a + panel_b + panel_c) / 
#   (panel_d + (panel_e/panel_g) + (panel_f/panel_h) ) / 
#   (panel_i1 + panel_i2 + panel_j + panel_k + plot_layout(widths = c(1,1,1.5,1.5))) + plot_layout(heights = c(1,2,1.2))

# (panel_a + panel_b) / 
#   (panel_c + (panel_d/panel_f) + (panel_e/panel_g) ) / 
#   plot_spacer() /
#   (panel_h + panel_i + panel_j + plot_layout(widths = c(1.5,1,1))) + plot_layout(heights = c(0.9, 2,0.05, 1.2))

(panel_a + panel_b) / 
  (panel_c + ((panel_d +panel_e+ panel_f)/(panel_meta_g + panel_meta_h) ) + plot_layout(widths = c(1,4)) ) / 
  plot_spacer() /
  (panel_i + panel_j + panel_k + plot_layout(widths = c(1.5,1,1))) + plot_layout(heights = c(0.8, 2,0.05, 1.2))
```
## NEW EXPLORATION
## Try osc p-value statistics
#### (j) pctm at DNA synthesis
```{r fig.height = 5,fig.width =8}

oscmodc <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  # filter(EXvivo_ENR_xCOLL_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  mutate(direction = ifelse(EXvivo_ENR_xCOLL_est < 0, "linear loss", "linear gain")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark",
                          MHR_ENR_xCOLL_pval > 0.05 ~ "not oscillating")) %>%
  filter(media == "ENR") %>%
  select(modid, zt, ct, pctm, age, acro, direction) %>%
  group_by(modid, age, acro, direction) %>%
  mutate(center = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm-center) %>%
  filter(zt == 16) %>%
  mutate(zt = ifelse(zt == 16, "peak S phase (~ZT16)", "peak G2/M (~ZT23)")) %>%
  mutate(cat = paste0(acro," : ",direction)) %>%
  filter(cat == "dark : linear gain" | cat == "light : linear loss")

q_pvals <-
  dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  group_by(acro) %>%
  select(modid, acro, MHR_ENR_xCOLL_pval) %>%
  # summarise(n())
  mutate(qbins = cut(MHR_ENR_xCOLL_pval, quantile(MHR_ENR_xCOLL_pval, seq(0,1,0.1)))) %>%
  rename(osc_pval = MHR_ENR_xCOLL_pval)
  # top_n(., 10, -MHR_ENR_xCOLL_pval)

oscmodc %>%
  left_join(., q_pvals) %>%
  # mutate(qbins = quantile(MHR_ENR_xCOLL_pval))
  # filter(modid %in% top_pvals$modid) %>%
  ggplot(aes(y=qbins , x=(pctm_norm)*100)) +
  geom_boxplot() +
  xlab("%modC @ peak S phase") +
  facet_wrap(~cat, scales="free")

oscmodc  %>%
  left_join(., q_pvals) %>%
  ggplot(aes(y=pctm_norm*100, x=-log10(osc_pval), col=cat)) +
  geom_point() +
  ylab("%modC @ peak S phase") +
  geom_smooth(method="lm")
```
## Try osc amplitude
#### (j) pctm at DNA synthesis
```{r fig.height = 5,fig.width =8}

oscmodc <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  # filter(EXvivo_ENR_xCOLL_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  mutate(direction = ifelse(EXvivo_ENR_xCOLL_est < 0, "linear loss", "linear gain")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark",
                          MHR_ENR_xCOLL_pval > 0.05 ~ "not oscillating")) %>%
  filter(media == "ENR") %>%
  select(modid, zt, ct, pctm, age, acro, direction) %>%
  group_by(modid, age, acro, direction) %>%
  mutate(center = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm-center) %>%
  filter(zt == 16) %>%
  mutate(zt = ifelse(zt == 16, "peak S phase (~ZT16)", "peak G2/M (~ZT23)")) %>%
  mutate(cat = paste0(acro," : ",direction)) %>%
  filter(cat == "dark : linear gain" | cat == "light : linear loss")

q_pvals <-
  dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  group_by(acro) %>%
  select(modid, acro, MHR_ENR_xCOLL_amp) %>%
  # summarise(n())
  mutate(qbins = cut(MHR_ENR_xCOLL_amp, quantile(MHR_ENR_xCOLL_amp, seq(0,1,0.1)), include.lowest=T)) %>%
  rename(osc_amp = MHR_ENR_xCOLL_amp)
  # top_n(., 10, -MHR_ENR_xCOLL_pval)

oscmodc %>%
  left_join(., q_pvals) %>%
  # mutate(qbins = quantile(MHR_ENR_xCOLL_pval))
  # filter(modid %in% top_pvals$modid) %>%
  ggplot(aes(y=qbins , x=(pctm_norm)*100)) +
  geom_boxplot() +
  xlab("%modC @ peak S phase") +
  facet_wrap(~cat, scales="free")

oscmodc  %>%
  left_join(., q_pvals) %>%
  ggplot(aes(y=pctm_norm*100, x=(osc_amp), col=cat)) +
  geom_point() +
  ylab("%modC @ peak S phase") +
  geom_smooth(method="lm")
```

## Try osc acrophase
#### (j) pctm at DNA synthesis
```{r fig.height = 3,fig.width =5}

oscmodc <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  # filter(EXvivo_ENR_xCOLL_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  mutate(direction = ifelse(EXvivo_ENR_xCOLL_est < 0, "linear loss", "linear gain")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark",
                          MHR_ENR_xCOLL_pval > 0.05 ~ "not oscillating")) %>%
  filter(media == "ENR") %>%
  select(modid, zt, ct, pctm, age, acro, direction) %>%
  group_by(modid, age, acro, direction) %>%
  mutate(center = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm-center) %>%
  filter(zt == 16) %>%
  mutate(zt = ifelse(zt == 16, "peak S phase (~ZT16)", "peak G2/M (~ZT23)")) %>%
  mutate(cat = paste0(acro," : ",direction)) %>%
  filter(cat == "dark : linear gain" | cat == "light : linear loss")

q_acro <-
  dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  group_by(acro) %>%
  select(modid, acro, MHR_ENR_xCOLL_acro) %>%
  # summarise(n())
  mutate(qbins = cut(MHR_ENR_xCOLL_acro, seq(0,24,3), include.lowest=T)) %>%
  rename(osc_acro = MHR_ENR_xCOLL_acro)
  # top_n(., 10, -MHR_ENR_xCOLL_pval)

oscmodc %>%
  left_join(., q_acro) %>%
  # mutate(qbins = quantile(MHR_ENR_xCOLL_pval))
  # filter(modid %in% top_pvals$modid) %>%
  ggplot(aes(x=qbins , y=(pctm_norm)*100)) +
  geom_boxplot() 

oscmodc  %>%
  left_join(., q_acro) %>%
  ggplot(aes(y=pctm_norm*100, x=osc_acro, col=qbins)) +
  ylab("%modC @ peak S phase") +
  geom_boxplot() +
  geom_point() 

q_acro %>%
  group_by(qbins) %>%
  summarise(n= n())
```

## Try best agers pvals
#### (j) pctm at DNA synthesis
```{r fig.height = 3,fig.width =5}

oscmodc <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  # filter(EXvivo_ENR_xCOLL_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  mutate(direction = ifelse(EXvivo_ENR_xCOLL_est < 0, "linear loss", "linear gain")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark",
                          MHR_ENR_xCOLL_pval > 0.05 ~ "not oscillating")) %>%
  filter(media == "ENR") %>%
  select(modid, zt, ct, pctm, age, acro, direction) %>%
  group_by(modid, age, acro, direction) %>%
  mutate(center = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm-center) %>%
  filter(zt == 16) %>%
  mutate(zt = ifelse(zt == 16, "peak S phase (~ZT16)", "peak G2/M (~ZT23)")) %>%
  mutate(cat = paste0(acro," : ",direction)) %>%
  filter(cat == "dark : linear gain" | cat == "light : linear loss")

q_pvals <-
  dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  group_by(acro) %>%
  select(modid, acro, EXvivo_ENR_xCOLL_pval) %>%
  # summarise(n())
  mutate(qbins = cut(EXvivo_ENR_xCOLL_pval, quantile(EXvivo_ENR_xCOLL_pval, seq(0,1,0.1)), include.lowest=T)) %>%
  rename(exvivo_pval = EXvivo_ENR_xCOLL_pval)
  # top_n(., 10, -MHR_ENR_xCOLL_pval)

oscmodc %>%
  left_join(., q_pvals) %>%
  # mutate(qbins = quantile(MHR_ENR_xCOLL_pval))
  # filter(modid %in% top_pvals$modid) %>%
  ggplot(aes(y=qbins , x=(pctm_norm)*100)) +
  geom_boxplot() +
  xlab("%modC @ peak S phase") +
  facet_wrap(~cat, scales="free")

oscmodc  %>%
  left_join(., q_pvals) %>%
  ggplot(aes(y=pctm_norm*100, x=-log10(exvivo_pval), col=cat)) +
  geom_point() +
  ylab("%modC @ peak S phase") +
  geom_smooth(method="lm")
```
## Try best agers estimate
#### (j) pctm at DNA synthesis
```{r fig.height = 3,fig.width =5}

oscmodc <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  # filter(EXvivo_ENR_xCOLL_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  mutate(direction = ifelse(EXvivo_ENR_xCOLL_est < 0, "linear loss", "linear gain")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark",
                          MHR_ENR_xCOLL_pval > 0.05 ~ "not oscillating")) %>%
  filter(media == "ENR") %>%
  select(modid, zt, ct, pctm, age, acro, direction) %>%
  group_by(modid, age, acro, direction) %>%
  mutate(center = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm-center) %>%
  filter(zt == 16) %>%
  mutate(zt = ifelse(zt == 16, "peak S phase (~ZT16)", "peak G2/M (~ZT23)")) %>%
  mutate(cat = paste0(acro," : ",direction)) %>%
  filter(cat == "dark : linear gain" | cat == "light : linear loss")

q_pvals <-
  dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  group_by(acro) %>%
  select(modid, acro, EXvivo_ENR_xCOLL_est) %>%
  # summarise(n())
  mutate(qbins = cut(abs(EXvivo_ENR_xCOLL_est), quantile(abs(EXvivo_ENR_xCOLL_est), seq(0,1,0.1)), include.lowest=T)) %>%
  rename(exvivo_est = EXvivo_ENR_xCOLL_est)
  # top_n(., 10, -MHR_ENR_xCOLL_pval)

oscmodc %>%
  left_join(., q_pvals) %>%
  # mutate(qbins = quantile(MHR_ENR_xCOLL_pval))
  # filter(modid %in% top_pvals$modid) %>%
  ggplot(aes(y=qbins , x=(pctm_norm)*100)) +
  geom_boxplot() +
  xlab("%modC @ peak S phase") +
  facet_wrap(~cat, scales="free")

oscmodc  %>%
  left_join(., q_pvals) %>%
  ggplot(aes(y=pctm_norm*100, x=abs(exvivo_est), col=cat)) +
  geom_point() +
  ylab("%modC @ peak S phase") +
  geom_smooth(method="lm")
```

### end
#### (i) osc-modCs: Significant only 
```{r fig.height = 2.5,fig.width =5}

oscmodc <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05) %>%
  # filter(MHR_ENR_xCOLL_pval < 0.05 & HR_tl_s4UCORCT_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  # mutate(acro = ifelse(MHR_ENR_xCOLL_acro < 12, "osc-modC\nacrophase\n(day)",
  #                      "osc-modC\nacrophase\n(night)")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  select(modid, zt, ct, pctm, age, media, acro) %>%
  filter(media == "ENR") %>%
  group_by(modid, media, age) %>%
  mutate(pctm_age = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm - pctm_age) 

top_pvals <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05) %>%
    mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  group_by(acro) %>%
  top_n(., 10, -MHR_ENR_xCOLL_pval)


```

```{r fig.height = 2.5,fig.width =4}

# panel_i <-
  oscmodc %>%
  # mutate(zt = zt + 24) %>%
  # rbind(., oscmodc) %>%
  # mutate(zt = zt + 24) %>%
  # rbind(., oscmodc) %>%
  filter(between(zt, 0, 24)) %>%
  filter(zt <= 24) %>%
  filter(modid %in% top_pvals$modid) %>%
  # filter(acro == "(3,6]" | acro == "(6,9]" | acro == "(9,12]"| acro == "(18,21]" |acro == "(21,24]") %>%
  mutate(acro = factor(acro, levels=c("light", "dark", "NA"))) %>%
  # filter(acro == "acrophase:\nZT:0-12") %>%
  filter(!is.na(acro)) %>%
  ggplot(aes(x=zt, y=pctm_norm*100)) +
  annotate("rect",ymin=-Inf, ymax=Inf, xmin = 13, xmax=19, fill="grey", alpha=0.5) +
  # facet_wrap(acro~., nrow=1, strip.position = "top")  +
  theme(legend.position = "none", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  # geom_vline(xintercept=c(S_acro, S_acro+24), linetype="dotted", col="black") +
  geom_smooth(aes(group=as.factor(modid), col=acro),
              # col="#999999",
              method="lm", se=F, size=0.25, alpha=0.25,
              formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi), fullrange=TRUE) +
  # annotate("rect",ymin=-Inf, ymax=Inf,
  #          xmin = G2M_acro-G2M_acro_sd-24, xmax=G2M_acro+G2M_acro_sd-24, fill="#980043", alpha=0.1) +
  # annotate("rect",ymin=-Inf, ymax=Inf, 
  #          xmin = G2M_acro-G2M_acro_sd, xmax=G2M_acro+G2M_acro_sd, fill="#980043", alpha=0.1) +
  # annotate("rect",ymin=-Inf, ymax=Inf, 
  #          xmin = G2M_acro-G2M_acro_sd+24, xmax=G2M_acro+G2M_acro_sd+24, fill="#980043", alpha=0.1) +
  # geom_vline(xintercept=c(G2M_acro, G2M_acro+24, G2M_acro-24), linetype="dashed", col="#980043") +
  # annotate("rect",ymin=-Inf, ymax=Inf, xmin = 13+24, xmax=19+24, fill="grey", alpha=0.5) +
  # annotate("rect",ymin=-Inf, ymax=Inf, xmin = 10+48, xmax=22+48, fill="blue", alpha=0.2) +
  # annotate("rect",ymin=-Inf, ymax=Inf, xmin = S_acro-S_acro_sd+24, xmax=S_acro+S_acro_sd+24, fill="blue", alpha=0.2) +
  geom_smooth(aes(col=acro), method="lm", se=F, size=2, alpha=0.1, 
              formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi), fullrange=TRUE) +
  geom_vline(xintercept=c(S_acro), linetype="dotted", col="black") +
  ylab("normalized % modC") +
  xlab("ZT") +
  # ylim(-2.5,2.5) +
  # scale_color_manual(values=c(viridis::viridis(1000))) +
  # scale_color_grey() +
  scale_color_manual(values=c("darkred","darkblue")) +
  # scale_x_continuous(breaks=seq(0, 24, 4), labels=c(rep(seq(0, 20, 4),1),0))  + 
  ggtitle("i") + 
  theme(strip.background =element_blank(),
        text = element_text(size = 7))
```

#### (j) pctm at DNA synthesis
```{r fig.height = 5,fig.width =8}

oscmodc <- dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  # filter(EXvivo_ENR_xCOLL_pval < 0.05) %>%
  fetch_pctm(., mydat) %>%
  mutate(direction = ifelse(EXvivo_ENR_xCOLL_est < 0, "linear loss", "linear gain")) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_pval < 0.05 & MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark",
                          MHR_ENR_xCOLL_pval > 0.05 ~ "not oscillating")) %>%
  filter(media == "ENR") %>%
  select(modid, zt, ct, pctm, age, acro, direction) %>%
  group_by(modid, age, acro, direction) %>%
  mutate(center = median(pctm)) %>%
  ungroup() %>%
  mutate(pctm_norm = pctm-center) %>%
  filter(zt == 16) %>%
  mutate(zt = ifelse(zt == 16, "peak S phase (~ZT16)", "peak G2/M (~ZT23)")) %>%
  mutate(cat = paste0(acro," : ",direction)) %>%
  filter(cat == "dark : linear gain" | cat == "light : linear loss")

q_pvals <-
  dat %>%
  filter(MHR_ENR_xCOLL_pval < 0.05 & EXvivo_ENR_xCOLL_pval <0.05) %>%
  mutate(acro = case_when(MHR_ENR_xCOLL_acro > 0 & MHR_ENR_xCOLL_acro < 12~"light",
                          MHR_ENR_xCOLL_acro > 12 & MHR_ENR_xCOLL_acro < 24~"dark")) %>%
  group_by(acro) %>%
  select(modid, acro, MHR_ENR_xCOLL_pval) %>%
  # summarise(n())
  mutate(qbins = cut(MHR_ENR_xCOLL_pval, quantile(MHR_ENR_xCOLL_pval, seq(0,1,0.1)))) %>%
  rename(osc_pval = MHR_ENR_xCOLL_pval)
  # top_n(., 10, -MHR_ENR_xCOLL_pval)

oscmodc %>%
  left_join(., q_pvals) %>%
  # mutate(qbins = quantile(MHR_ENR_xCOLL_pval))
  # filter(modid %in% top_pvals$modid) %>%
  ggplot(aes(y=qbins , x=(pctm_norm)*100)) +
  geom_boxplot() +
  facet_wrap(~cat, scales="free")

oscmodc  %>%
  left_join(., q_pvals) %>%
  ggplot(aes(y=pctm_norm*100, x=-log10(osc_pval), col=cat)) +
  geom_point() +
  geom_smooth(method="lm")
```

```{r}

```
