---
title: Core Findings
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
---

- Goal of this page is to summarize cytosine modification data.
- I forwent EVC usage. It makes no difference for this noisy dataset.
 
 
# Import data

```{r}
require(data.table)
require(matrixStats)
require(dplyr)
require(tidyr)
require(ggplot2)
require(annmatrix)
# source("code/environmentSetup.R")
source("code/fns.R")
# devtools::load_all("rsrc")
sincos2acro <- function(sin,cos, per=24) (atan2(sin,cos)/2/pi*per+per)%%per

sincos2amp <- function(sin,cos) 2*sqrt(sin^2+cos^2)

### Files from Matt
pctm <- readRDS("output/ReduceOutliers/mydat_multirep.RDS")
# aovres <- readRDS("output/ReliableCytosines.RDS")
pctm <- pctm[,pctm$age != 20]

pctm@pos <- pctm@position
pctm@modid <- paste(pctm@chr, pctm@pos, pctm@strand, sep=":")
# pctm@aov_pval <- aovres$pvalue

pctm$sin <- sin(pctm$ct/24*2*pi)
pctm$cos <- cos(pctm$ct/24*2*pi)
pctm$age <- as.numeric(as.character(pctm$age))

### Fix ZT and time of day; continuous time (ct) does not change
### ZT NOT USED for any calculation. It's just for plotting, but I thought I'd fix it early.
pctm$timeofday <- pctm$zt
pctm$zt <- ct2zt_e3e4(pctm$ct)

###
pctm$batch <- case_when(pctm$age == 16 ~ "B1", pctm$age == 31 ~ "B2",
          pctm$age == 62 ~ "B1")

```


## FACS data {.tabset}

### Temporal plots {.tabset}
#### By CT
```{r fig.height=2.5, fig.width=7}
facs <- fread("input/flow_analysis.csv")
facs$zt <- ct2zt(facs$ct)

g1 <- facs %>%
  subset(Var1 == "G1" ) %>%
  ggplot(aes(ct, perCount, color=Var1))+
  geom_point(alpha=0.5)+
  # geom_smooth(,method="loess", span=0.5,se=F)+
  geom_smooth(method="lm", se=T, alpha=0.3) + 
  scale_x_continuous(breaks=seq(0,24,3)) +
  scale_color_manual(values=c("#dd1c77")) + ylab("cell proportion (%)") +xlab("continuous time (CT)")

g2 <- facs %>%
  subset(Var1 == "G2/M" | Var1 == "S") %>%
  ggplot(aes(ct, perCount, color=Var1))+
  geom_point(alpha=0.5)+
  # geom_smooth(,method="loess", span=0.5,se=F)+
  geom_smooth(method="lm", se=T, alpha=0.3,formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi)) + 
  scale_x_continuous(breaks=seq(0,24,3)) +
  scale_color_manual(values=c("#999999", "#980043")) + ylab("cell proportion (%)")+xlab("continuous time (CT)")
  
gridExtra::grid.arrange(g2, g1, ncol=2)

```

#### By ZT
```{r fig.height=3, fig.width=4}

facs %>%
  subset(Var1 == "G2/M" | Var1 == "S") %>%
  ggplot(aes(zt, perCount, color=Var1))+
  geom_point(alpha=0.5)+
  # geom_smooth(,method="loess", span=0.5,se=F)+
  geom_smooth(method="lm", se=T, alpha=0.3,formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi)) + 
  scale_x_continuous(breaks=seq(0,24,3)) +
  scale_color_manual(values=c("#999999", "#980043")) + ylab("cell proportion (%)")+xlab("zeitgeber time (ZT)")
  
```

### Harmonic regressionZT
#### G2/M 
```{r fig.height=3, fig.width=6}

test <- facs %>%
  select(Var1, ct, zt, rep, perCount) %>%
  mutate(Var1 = gsub("/","_",Var1)) %>%
  spread(., key = "Var1", value=perCount)

res <- lm(G2_M ~ sin(ct/24*2*pi) + cos(ct/24*2*pi) + ct, data=test)
res2 <- lm(G2_M ~ ct, data=test)

anova(res, res2)

# sincos2acro(sin=res$coefficients[[2]], cos=res$coefficients[[3]])

tmp1 <- lm(G2_M ~  sin(zt/24*2*pi) + cos(zt/24*2*pi), data=test %>% filter(rep=="A"))
tmp2 <- lm(G2_M ~  sin(zt/24*2*pi) + cos(zt/24*2*pi), data=test %>% filter(rep=="B"))
tmp3 <- lm(G2_M ~  sin(zt/24*2*pi) + cos(zt/24*2*pi), data=test %>% filter(rep=="C"))

G2_M_acro <- rbind(sincos2acro(tmp1$coefficients[[2]], tmp1$coefficients[[3]]),
              sincos2acro(tmp2$coefficients[[2]], tmp2$coefficients[[3]]),
              sincos2acro(tmp3$coefficients[[2]], tmp3$coefficients[[3]]))

24 + mean(ifelse(G2_M_acro >12, G2_M_acro - 24, G2_M_acro))
sd(ifelse(G2_M_acro >12, G2_M_acro - 24, G2_M_acro))

```
#### S 
```{r fig.height=3, fig.width=6}

res <- lm(S ~ sin(ct/24*2*pi) + cos(ct/24*2*pi) + ct, data=test)
res2 <- lm(S ~ ct, data=test)

anova(res, res2)

tmp1 <- lm(S ~  sin(zt/24*2*pi) + cos(zt/24*2*pi), data=test %>% filter(rep=="A"))
tmp2 <- lm(S ~  sin(zt/24*2*pi) + cos(zt/24*2*pi), data=test %>% filter(rep=="B"))
tmp3 <- lm(S ~  sin(zt/24*2*pi) + cos(zt/24*2*pi), data=test %>% filter(rep=="C"))

S_acro <- rbind(sincos2acro(tmp1$coefficients[[2]], tmp1$coefficients[[3]]),
              sincos2acro(tmp2$coefficients[[2]], tmp2$coefficients[[3]]),
              sincos2acro(tmp3$coefficients[[2]], tmp3$coefficients[[3]]))
mean(S_acro)
sd(S_acro)

```

#### G1
```{r fig.height=3, fig.width=6}

res <- lm(G1 ~ sin(ct/24*2*pi) + cos(ct/24*2*pi) + ct, data=test)
res2 <- lm(G1 ~ ct, data=test)

anova(res, res2)

```


### Imputing values {.tabset}
```{r fig.height=3, fig.width=6}
facs %>%
  subset(Var1 == "G2/M" |Var1 =="G1" | Var1 == "S") %>%
  group_by(Var1, ct) %>%
  summarise(mean = median(perCount))  %>%
  ggplot(aes(ct, mean)) +
  geom_point() +
  geom_smooth(method="lm", se=T, alpha=0.3,formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi)) + 
  facet_wrap(~Var1, scales="free")

mean_facs <- facs %>%
  group_by(Var1, ct) %>%
  summarise(mean = mean(perCount)) %>%
  spread(., key=Var1, value=mean)

```

#### G1
```{r fig.height=3, fig.width=6}
res <- lm(G1~ct, data=mean_facs)
summary(res)
mean_facs$G1_imputed <- sapply(1:nrow(mean_facs), function(x) res$coefficients[2]*mean_facs$ct[x] + res$coefficients[1])

mean_facs %>%
  select(ct, G1, G1_imputed) %>%
  melt(., id.var="ct") %>%
  ggplot(aes(ct, value, col=variable)) +
  geom_point() +
  geom_smooth(method="lm")

Imputed_values <- data.frame(ct=0:32, G1_imputed = sapply(0:32, function(x) res$coefficients[2]*x + res$coefficients[1]))
```

#### G2/M 
```{r fig.height=3, fig.width=6}
res <- lm(`G2/M` ~  sin(ct/24*2*pi) + cos(ct/24*2*pi), data=mean_facs)
res$coefficients
mean_facs$G2M_imputed <- sapply(1:nrow(mean_facs), 
                                function(x) res$coefficients[2]*sin(mean_facs$ct[x]/24*2*pi) +
                                  res$coefficients[3]*cos(mean_facs$ct[x]/24*2*pi) + 
                                  res$coefficients[1])


mean_facs %>%
  select(ct, `G2/M`, G2M_imputed) %>%
  melt(., id.var="ct") %>%
  ggplot(aes(ct, value, col=variable)) +
  geom_point() +
  geom_smooth(method="lm", se=T, alpha=0.3,formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi))


Imputed_values$G2M_imputed <- sapply(0:32, 
                                function(x) res$coefficients[2]*sin(x/24*2*pi) +
                                  res$coefficients[3]*cos(x/24*2*pi) + 
                                  res$coefficients[1])

```
#### S 
```{r fig.height=3, fig.width=6}
res <- lm(S ~  sin(ct/24*2*pi) + cos(ct/24*2*pi), data=mean_facs)
res$coefficients
mean_facs$S_imputed <- sapply(1:nrow(mean_facs), 
                                function(x) res$coefficients[2]*sin(mean_facs$ct[x]/24*2*pi) +
                                  res$coefficients[3]*cos(mean_facs$ct[x]/24*2*pi) + 
                                  res$coefficients[1])


mean_facs %>%
  select(ct, S, S_imputed) %>%
  melt(., id.var="ct") %>%
  ggplot(aes(ct, value, col=variable)) +
  geom_point() +
  geom_smooth(method="lm", se=T, alpha=0.3,formula = y ~ sin(x/24*2*pi) + cos(x/24*2*pi))


Imputed_values$S_imputed <- sapply(0:32, 
                                function(x) res$coefficients[2]*sin(x/24*2*pi) +
                                  res$coefficients[3]*cos(x/24*2*pi) + 
                                  res$coefficients[1])
```

## Add FACS data to pctm ss
```{r}

IV <- data.frame(ct= pctm$ct, pipelinerunid = pctm$pipelinerunid) %>%
  left_join(., Imputed_values, by="ct")

all(IV$pipelinerunid == pctm$pipelinerunid)
all(IV$pipelinerunid == colnames(pctm))

pctm$G2M_im <- IV$G2M_imputed
pctm$G1_im <- IV$G1_imputed
pctm$S_im <- IV$S_imputed

```

```{r}
# Prep data
mydat <- avgBioTech(pctm)

# saveRDS(pctm, "output/ReduceOutliers/mydat_multirep_forperm.RDS")
# mydat <- pctm

# Used for temporal variation only
mydat_biotech <- pctm


```


# Detecting Biological Effects

## Ex-vivo Drifting Effects

Linear time effects on a short time scale (~1 day of a mouse's ~700 day? lifespan) in the organoid system could be observed:  
a) If there exists a large aging effect. Still unlikely if uniform linear over 700 days (1 day gives 1/700 = 0.14%/day)  
b) If an aging effect is moderate and specific to a short developmental period  
c) If organoids "accelerate" an aging effect  
d) If the organoid system has age-unrelated "temporal drifts"  

Similarly to the technical vs inter-organoid/time hypothesis test, here it will be tested if it is likely that all organoids are similar given a more specific alternative hypothesis (linear time). Since there are two alternative hypotheses (oscillation, linear time), to test linear time it will be assumed oscillations exist and vice versa (type II ANOVA).

Null:  `~sin(ct) + cos(ct)`  
Alt: `~sin(ct) + cos(ct) + ct`  

<!-- ### By media and age -->

<!-- ```{r fig.height=3, fig.width=6} -->
<!-- CTeff <- annapply(mydat,c("media","age"),function(x){ -->
<!-- 		fit <- annOLSwFtest(x,~sin+cos+ct,~sin+cos,ver=T) -->
<!-- 		data.frame("value"=fit$pvalue,fit$fit$betas, rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("media","age"), remove=FALSE) -->

<!-- CTeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(media~age)+ -->
<!-- 	xlab("CT P-value") -->

<!-- CTeff %>%  -->
<!-- 	group_by(age,media) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~ age + media)  -->

<!-- pval.hr <- CTeff %>% -->
<!--   select(modid, grp, value) %>% -->
<!--   spread(.,  key = grp, value=value) %>% -->
<!--   arrange(modid) -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("EXvivo", colnames(pval.hr), "pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- acro.hr <- CTeff %>% -->
<!--   select(modid, grp, ct) %>% -->
<!--   spread(.,  key = grp, value=ct) -->

<!-- acro.hr <- acro.hr[match(mydat@modid, acro.hr$modid),] -->
<!-- colnames(acro.hr) <- paste("EXvivo", colnames(acro.hr), "est", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), acro.hr[,2:ncol(acro.hr)]) -->

<!-- ``` -->

### By media
```{r fig.height=3, fig.width=6}
CTeff <- annapply(mydat,c("media"),function(x){
		fit <- annOLSwFtest(x,~sin+cos+ct+as.factor(age),~sin+cos+as.factor(age),ver=T)
		data.frame("value"=fit$pvalue,fit$fit$betas, rowanns(x))
	})

CTeff %>%
	ggplot(aes(value))+
	geom_histogram(breaks=0:20/20)+
	facet_grid(.~grp)+
	xlab("CT P-value")

CTeff %>% 
	group_by(grp) %>% 
	summarize(res=list(summaryPvalTable(value))) %>% 
  tidyr::unnest(res) %>%
	dcast(level ~ grp) 

pval.hr <- CTeff %>%
  select(modid, grp, value) %>%
  spread(.,  key = grp, value=value) %>%
  arrange(modid)

pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),]
colnames(pval.hr) <- paste("EXvivo", colnames(pval.hr), "xCOLL_pval", sep="_")
mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)])

acro.hr <- CTeff %>%
  select(modid, grp, ct) %>%
  spread(.,  key = grp, value=ct)

acro.hr <- acro.hr[match(mydat@modid, acro.hr$modid),]
colnames(acro.hr) <- paste("EXvivo", colnames(acro.hr), "xCOLL_est", sep="_")
mydat@'' <- cbind(rowanns(mydat), acro.hr[,2:ncol(acro.hr)])

```

<!-- ## Temporal effects {.tabset} -->

<!-- Null:  `~as.factor(ct)`   -->
<!-- Alt: `~1`   -->

<!-- ### media and age -->

<!-- ```{r fig.height=3, fig.width=6} -->

<!-- annapply_fill <- function(x,cols,fun,comb=T){ -->
<!--   grp <- apply(colanns(x)[,cols,drop=F],1,paste,collapse="_") -->
<!--   res <- lapply(csplit(x,grp),fun)  -->

<!--   # Tries to generate a data.frame -->
<!--   if(comb){ -->
<!--     if("data.frame" %in% class(res[[1]])){ -->
<!--       return(res %>% data.table::rbindlist(idcol="grp", fill=TRUE)) -->
<!--     } else if(is.vector(res[[1]])) { -->
<!--       return(lapply(res,function(x) data.frame("value"=x)) %>% data.table::rbindlist(idcol="grp", fill=TRUE)) -->
<!--     }else { -->
<!--       warning("Could not aggregate results, returning a list") -->
<!--       return(res) -->
<!--     } -->
<!--   } else { -->
<!--     return(res) -->
<!--   } -->
<!-- } -->

<!-- TEMeff <- annapply_fill(mydat_biotech, c("media","age"), function(x){ -->
<!--     res <-  rowVars(x, na.rm = TRUE) -->
<!-- 		fit <- annOLSwFtest(x, ~as.factor(ct), ~1, ver=T) -->
<!-- 		data.frame("variance"=res, -->
<!-- 		           "value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("media", "age"), remove=FALSE) -->

<!-- TEMeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(media~age)+ -->
<!-- 	xlab("temporal P-value") -->

<!-- TEMeff %>%  -->
<!-- 	group_by(age,media) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~ age + media)  -->

<!-- pval.hr <- TEMeff %>% -->
<!--   select(modid, grp, value) %>% -->
<!--   spread(.,  key = grp, value=value)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MTEM", colnames(pval.hr), "pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- acro.hr <- TEMeff %>% -->
<!--   select(modid, grp, variance) %>% -->
<!--   spread(.,  key = grp, value=variance) -->

<!-- acro.hr <- acro.hr[match(mydat@modid, acro.hr$modid),] -->
<!-- colnames(acro.hr) <- paste("MVAR", colnames(acro.hr), "var", sep="_") -->
<!-- all(acro.hr$MVAR_modid_var == mydat@modid) -->
<!-- mydat@'' <- cbind(rowanns(mydat), acro.hr[,2:ncol(acro.hr)]) -->

<!-- ``` -->

<!-- ### media -->

<!-- ```{r fig.height=3, fig.width=6} -->

<!-- TEMeff <- annapply_fill(mydat, c("media"), function(x){ -->
<!--     res <-  rowVars(x, na.rm = TRUE) -->
<!-- 		fit <- annOLSwFtest(x, ~as.factor(ct) + as.factor(age), ~as.factor(age), ver=T) -->
<!-- 		data.frame("variance"=res, -->
<!-- 		           "value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("media"), remove=FALSE) -->

<!-- TEMeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(media~.)+ -->
<!-- 	xlab("temporal P-value") -->

<!-- TEMeff %>%  -->
<!-- 	group_by(media) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~  media)  -->

<!-- pval.hr <- TEMeff %>% -->
<!--   select(modid, grp, value) %>% -->
<!--   spread(.,  key = grp, value=value)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MTEM", colnames(pval.hr), "xCOLL_pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- acro.hr <- TEMeff %>% -->
<!--   select(modid, grp, variance) %>% -->
<!--   spread(.,  key = grp, value=variance) -->

<!-- acro.hr <- acro.hr[match(mydat@modid, acro.hr$modid),] -->
<!-- colnames(acro.hr) <- paste("MVAR", colnames(acro.hr), "xCOLL_var", sep="_") -->
<!-- all(acro.hr$MVAR_modid_var == mydat@modid) -->
<!-- mydat@'' <- cbind(rowanns(mydat), acro.hr[,2:ncol(acro.hr)]) -->


<!-- ``` -->

## Oscillation Effects  {.tabset}

<!-- ### Media and age -->
<!-- ```{r fig.height=3, fig.width=6} -->
<!-- HReff <- annapply(mydat,c("media","age"),function(x) { -->
<!-- 		fit <- annOLSwFtest(x,~sin+cos+ct,~ ct,ver=T) -->
<!-- 		data.frame("value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("media","age"), remove=FALSE) %>%  -->
<!--   mutate(acro=sincos2acro(sin,cos)) %>% -->
<!--   mutate(amp=sincos2amp(sin,cos)) -->

<!-- HReff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(media~age)+ -->
<!-- 	xlab("HR P-value") -->

<!-- HReff %>%  -->
<!-- 	group_by(age,media) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~ age + media)  -->


<!-- pval.hr <- HReff %>% -->
<!--   select(modid, grp, value) %>% -->
<!--   spread(.,  key = grp, value=value)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MHR", colnames(pval.hr), "pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- pval.hr <- HReff %>% -->
<!--   select(modid, grp, acro) %>% -->
<!--   spread(.,  key = grp, value=acro)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MHR", colnames(pval.hr), "acro", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- pval.hr <- HReff %>% -->
<!--   select(modid, grp, amp) %>% -->
<!--   spread(.,  key = grp, value=amp)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MHR", colnames(pval.hr), "amp", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- ``` -->


### Media 
```{r fig.height=3, fig.width=6}
HReff <- annapply(mydat,c("media"),function(x) {
		fit <- annOLSwFtest(x,~sin+cos+as.factor(age)*ct,~ as.factor(age)*ct,ver=T)
		data.frame("value"=fit$pvalue,
		           fit$fit$betas,
		           rowanns(x))
	}) %>%
  separate(grp, c("media"), remove=FALSE) %>% 
  mutate(acro=sincos2acro(sin,cos)) %>%
  mutate(amp=sincos2amp(sin,cos))

HReff %>%
  # filter(aov_pval < 0.1) %>%
	ggplot(aes(value))+
	geom_histogram(breaks=0:20/20)+
	facet_grid(~media)+
	xlab("HR P-value")

HReff %>% 
	group_by(media) %>% 
	summarize(res=list(summaryPvalTable(value))) %>% 
  tidyr::unnest(res) %>%
	dcast(level ~  media) 

pval.hr <- HReff %>%
  select(modid, grp, value) %>%
  spread(.,  key = grp, value=value) 

pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),]
colnames(pval.hr) <- paste("MHR", colnames(pval.hr), "xCOLL_pval", sep="_")
mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)])

pval.hr <- HReff %>%
  select(modid, grp, acro) %>%
  spread(.,  key = grp, value=acro) 

pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),]
colnames(pval.hr) <- paste("MHR", colnames(pval.hr), "xCOLL_acro", sep="_")
mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)])

pval.hr <- HReff %>%
  select(modid, grp, amp) %>%
  spread(.,  key = grp, value=amp) 

pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),]
colnames(pval.hr) <- paste("MHR", colnames(pval.hr), "xCOLL_amp", sep="_")
mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)])

```

<!-- ## ENRIV DMR {.tabset} -->

<!-- ### By age -->
<!-- ```{r fig.height=3, fig.width=6} -->
<!-- DMReff <- annapply(mydat,c("age"),function(x) { -->
<!-- 		fit <- annOLSwFtest(x, ~media + ct, ~ ct, ver=T) -->
<!-- 		data.frame("value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("age"), remove=FALSE)  -->

<!-- DMReff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(.~age)+ -->
<!-- 	xlab("IV DMR P-value") -->

<!-- DMReff %>%  -->
<!-- 	group_by(age) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res)  %>% -->
<!-- 	dcast(level ~ age) -->

<!-- pval.hr <- DMReff %>% -->
<!--   select(modid, grp, value) %>% -->
<!--   spread(.,  key = grp, value=value)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("IVDMR", colnames(pval.hr), "pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- pval.hr <- DMReff %>% -->
<!--   select(modid, grp, mediaENRIV) %>% -->
<!--   spread(.,  key = grp, value=mediaENRIV)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("IVDMR", colnames(pval.hr), "est", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- ``` -->

<!-- ### All -->
<!-- ```{r fig.height=3, fig.width=6} -->

<!-- mydat$test_all <- "test_all" -->
<!-- DMReff2 <- annapply(mydat,c("test_all"),function(x) { -->
<!-- 		fit <- annOLSwFtest(x, ~media + as.factor(ct) + as.factor(age), ~ as.factor(ct) + as.factor(age), ver=T) -->
<!-- 		data.frame("value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	})  -->

<!-- DMReff2 %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	xlab("IV DMR P-value") -->

<!-- DMReff2 %>%  -->
<!-- 	group_by(grp) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) -->

<!-- mydat@IVDMR_xCOLL_pval <- DMReff2$value -->
<!-- mydat@IVDMR_xCOLL_est <- DMReff2$mediaENRIV -->

<!-- ``` -->

<!-- ## Cell cycle  {.tabset} -->
<!-- ### By media {.tabset} -->
<!-- #### G1 -->
<!-- ```{r fig.height=3, fig.width=6} -->
<!-- CCeff <- annapply(mydat,c("media"),function(x) { -->
<!-- 		fit <- annOLSwFtest(x, ~G1_im + G2M_im + S_im + as.factor(age), ~ G2M_im + S_im +as.factor(age), ver=T) -->
<!-- 		data.frame("value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("media"), remove=FALSE)  -->

<!-- CCeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(.~media)+ -->
<!-- 	xlab("G1 P-value") -->

<!-- CCeff %>%  -->
<!-- 	group_by(media) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~  media)  -->

<!-- pval.hr <- CCeff %>% -->
<!--   select(modid, grp, value) %>% -->
<!--   spread(.,  key = grp, value=value)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MCC_G1", colnames(pval.hr), "xCOLL_pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- ``` -->

<!-- #### G2/M -->
<!-- ```{r fig.height=3, fig.width=6} -->
<!-- CCeff <- annapply(mydat,c("media"),function(x) { -->
<!-- 		fit <- annOLSwFtest(x, ~G1_im + G2M_im + S_im + as.factor(age), ~ G1_im + S_im + as.factor(age), ver=T) -->
<!-- 		data.frame("value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("media"), remove=FALSE) -->

<!-- CCeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(.~media)+ -->
<!-- 	xlab("G2/M P-value") -->

<!-- CCeff %>%  -->
<!-- 	group_by(media) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~  media)  -->

<!-- pval.hr <- CCeff %>% -->
<!--   select(modid, grp, value) %>% -->
<!--   spread(.,  key = grp, value=value)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MCC_G2M", colnames(pval.hr), "xCOLL_pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- # sincos2acro <- function(sin,cos, per=24) (atan2(sin,cos)/2/pi*per+per)%%per -->
<!-- # sin2acro <- function(x, per=24) (sin(x/24*2*pi))%%24 -->
<!-- #  -->
<!-- # sin2acro <- function(x, per=24) tan(sin(x)/2/pi*per+24)%%24 -->
<!-- #  -->
<!-- #  -->
<!-- # CCeff %>% -->
<!-- #   filter(value < 0.05) %>% -->
<!-- #   ggplot(aes(acro)) + -->
<!-- #   geom_histogram() + -->
<!-- #   facet_wrap(~media) -->

<!-- ``` -->

<!-- #### S -->
<!-- ```{r fig.height=3, fig.width=6} -->
<!-- CCeff <- annapply(mydat,c("media"),function(x) { -->
<!-- 		fit <- annOLSwFtest(x, ~G1_im + G2M_im + S_im + as.factor(age), ~ G1_im + G2M_im + as.factor(age), ver=T) -->
<!-- 		data.frame("value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("media"), remove=FALSE)  -->

<!-- CCeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(.~media)+ -->
<!-- 	xlab("S P-value") -->

<!-- CCeff %>%  -->
<!-- 	group_by(media) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~ media)  -->

<!-- pval.hr <- CCeff %>% -->
<!--   select(modid, grp, value) %>% -->
<!--   spread(.,  key = grp, value=value)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MCC_S", colnames(pval.hr), "xCOLL_pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- ``` -->

<!-- ### By media and age {.tabset} -->
<!-- #### G1 -->
<!-- ```{r fig.height=3, fig.width=6} -->
<!-- CCeff <- annapply(mydat,c("media","age"),function(x) { -->
<!-- 		fit <- annOLSwFtest(x, ~G1_im + G2M_im + S_im, ~ G2M_im + S_im, ver=T) -->
<!-- 		data.frame("value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("media","age"), remove=FALSE)  -->

<!-- CCeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(media~age)+ -->
<!-- 	xlab("G1 P-value") -->

<!-- CCeff %>%  -->
<!-- 	group_by(age, media) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~ age + media)  -->

<!-- pval.hr <- CCeff %>% -->
<!--   select(modid, age, media, value) %>% -->
<!--   dcast(modid ~ media + age) -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MCC_G1", colnames(pval.hr), "pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- ``` -->

<!-- #### G2/M -->
<!-- ```{r fig.height=3, fig.width=6} -->
<!-- CCeff <- annapply(mydat,c("media","age"),function(x) { -->
<!-- 		fit <- annOLSwFtest(x, ~G1_im + G2M_im + S_im, ~ G1_im + S_im, ver=T) -->
<!-- 		data.frame("value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("media","age"), remove=FALSE)  -->

<!-- CCeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(media~age)+ -->
<!-- 	xlab("G2/M P-value") -->

<!-- CCeff %>%  -->
<!-- 	group_by(media, age) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~  media + age)  -->

<!-- pval.hr <- CCeff %>% -->
<!--   select(modid, age, media, value) %>% -->
<!--   dcast(modid ~ media + age) -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MCC_G2M", colnames(pval.hr), "pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- ``` -->

<!-- #### S -->
<!-- ```{r fig.height=3, fig.width=6} -->
<!-- CCeff <- annapply(mydat,c("media","age"),function(x) { -->
<!-- 		fit <- annOLSwFtest(x, ~G1_im + G2M_im + S_im, ~ G1_im + G2M_im, ver=T) -->
<!-- 		data.frame("value"=fit$pvalue, -->
<!-- 		           fit$fit$betas, -->
<!-- 		           rowanns(x)) -->
<!-- 	}) %>% -->
<!--   separate(grp, c("media","age"), remove=FALSE)  -->

<!-- CCeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(media~age)+ -->
<!-- 	xlab("S P-value") -->

<!-- CCeff %>%  -->
<!-- 	group_by(media, age) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~ media + age)  -->

<!-- pval.hr <- CCeff %>% -->
<!--   select(modid, age, media, value) %>% -->
<!--   dcast(modid ~ media + age) -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MCC_S", colnames(pval.hr), "pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- ``` -->

## Pseudo-Aging Effects  {.tabset}

Organoid pools were collected 4.5 days after extraction from native tissue. No native tissue intestinal samples are currently available to estimate aging effects, therefore, in order to approximate aging effects from this data, we can make some assumptions:

1) Epigenetic state of the animal of origin is preserved in the organoids (t=0 organoids == native tissue).    
2) What happens during the 4.5 day growth period is one of:  
  a) Ex-vivo drift effects are linear and can be represented by the 4.5-6 day sampling window.  
  b) Ex-vivo drift effects are negligible
  c) Ex-vivo drift effects are non-linear but similar for each age/animal  

Many of these assumptions have no basis to rely on, but may be our best approximation given the data. Going forward we can choose to model each animals native tissue using:  
Assuming 2a - backwards extrapolated values OR    
Assuming 2b/c - average of organoids should represent aging slopes  

Note: pseudo-aging effects estimated in each media are non-independent since the same 4 animals are in each condition.

Here it will be tested if it is likely that all animals are likely the same given a more specific alternative hypothesis (linear time).

Null:  `~ age`
Alt: `~ 1`

```{r}
tmp <- annapply(mydat,c("media","age"),function(x){
  matrixStats::rowMedians(x)
},comb=F) %>% do.call(cbind,.)

mydatAA <- annmatrix(tmp,
					rann = rowanns(mydat),
					cann = colanns(mydat)[match(colnames(tmp),apply(colanns(mydat)[,c("media","age"),drop=F],1,paste,collapse="_")),])

```

Given the sampling is uneven, certain time points will have more "leverage" than others.

```{r fig.height=2,fig.width=4}
ages <- as.numeric(unique(mydatAA$age))
tmpmod <- lm(rnorm(length(ages)) ~ ages)
levs <- hatvalues(tmpmod)
names(levs) <- ages
qplot(ages,levs)+ylim(c(0,NA))+ylab("Leverage")
```

<!-- ### Pseudo-aging -->

<!-- ```{r fig.height=3, fig.width=6} -->

<!-- PAeff <- annapply(mydatAA,"media",function(x){ -->
<!-- 	fit <- annOLSwFtest(x,~age,~ 1,ver=T) -->
<!-- 	data.frame("value"=fit$pvalue,fit$fit$betas, rowanns(x)) -->
<!-- }) -->

<!-- PAeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(~grp)+ -->
<!-- 	xlab("Pseudo-Age P-value") -->

<!-- PAeff %>%  -->
<!-- 	group_by(grp) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~ grp) -->

<!-- pval.hr <- PAeff %>% -->
<!--   select(modid, grp, value) %>% -->
<!--   spread(.,  key = grp, value=value)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MPseudo", colnames(pval.hr), "pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- pval.hr <- PAeff %>% -->
<!--   select(modid, grp, age  ) %>% -->
<!--   spread(.,  key = grp, value=age  )  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("MPseudo", colnames(pval.hr), "est", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->

<!-- ``` -->

### Pseudo-aging2 
- correcting for CT and batch

```{r fig.height=3, fig.width=6}

PAeff <- annapply(mydat,"media",function(x){
	fit <- annOLSwFtest(x,~ age + as.factor(ct) + batch,~ as.factor(ct) + batch,ver=T)
	data.frame("value"=fit$pvalue,fit$fit$betas, rowanns(x))
})

PAeff %>%
	ggplot(aes(value))+
	geom_histogram(breaks=0:20/20)+
	facet_grid(~grp)+
	xlab("Pseudo-Age P-value")

PAeff %>% 
	group_by(grp) %>% 
	summarize(res=list(summaryPvalTable(value))) %>% 
  tidyr::unnest(res) %>%
	dcast(level ~ grp)

pval.hr <- PAeff %>%
  select(modid, grp, value) %>%
  spread(.,  key = grp, value=value) 

pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),]
colnames(pval.hr) <- paste("MPseudo2", colnames(pval.hr), "pval", sep="_")
mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)])

pval.hr <- PAeff %>%
  select(modid, grp, age  ) %>%
  spread(.,  key = grp, value=age  ) 

pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),]
colnames(pval.hr) <- paste("MPseudo2", colnames(pval.hr), "est", sep="_")
mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)])

```

<!-- ## Age DMR  -->
<!-- - using biotechs and correcting for CT -->

<!-- ```{r fig.height=3, fig.width=6} -->

<!-- PAeff <- annapply(mydat,"media",function(x){ -->
<!-- 	fit <- annOLSwFtest(x,~ as.factor(age) + as.factor(ct) + batch, ~as.factor(ct) +batch, ver=T) -->
<!-- 	data.frame("value"=fit$pvalue,fit$fit$betas, rowanns(x)) -->
<!-- }) -->

<!-- PAeff %>% -->
<!-- 	ggplot(aes(value))+ -->
<!-- 	geom_histogram(breaks=0:20/20)+ -->
<!-- 	facet_grid(~grp)+ -->
<!-- 	xlab("Age DMR P-value") -->

<!-- PAeff %>%  -->
<!-- 	group_by(grp) %>%  -->
<!-- 	summarize(res=list(summaryPvalTable(value))) %>%  -->
<!--   tidyr::unnest(res) %>% -->
<!-- 	dcast(level ~ grp) -->

<!-- pval.hr <- PAeff %>% -->
<!--   select(modid, grp, value) %>% -->
<!--   spread(.,  key = grp, value=value)  -->

<!-- pval.hr <- pval.hr[match(mydat@modid, pval.hr$modid),] -->
<!-- colnames(pval.hr) <- paste("DMR", colnames(pval.hr), "pval", sep="_") -->
<!-- mydat@'' <- cbind(rowanns(mydat), pval.hr[,2:ncol(pval.hr)]) -->


<!-- ``` -->




# Export data for later

```{r}
# require(readr)

### Convert ct acrophases to zt to be more applicable to biology
tmp <- rowanns(mydat) %>% mutate_if(., grepl("acro", colnames(.)), function(x) ct2zt_e3e4(x))

if(all(tmp$modid == mydat@modid)){
  mydat@'' <- tmp
} else {stop("error matching coordinates")}

### December 18th, 2019 
# mydat$old_zt <- mydat$zt
# mydat$zt <- ct2zt_e3e4(mydat$ct)

saveRDS(mydat, "output/methylation_stats.RDS")

```

### Quick plots {.tabset}

#### Osc-modC acrophase

```{r fig.height = 3,fig.width =4}
mydat <- readRDS("output/methylation_stats.RDS")
dat <- rowanns(mydat)

dat %>%
  # mutate(ageDMR = DMR_ENR_pval < 0.05) %>%
  filter(MHR_ENR_xCOLL_pval < 0.05) %>%
  ggplot(aes(x=MHR_ENR_xCOLL_acro)) +
    geom_rect(aes(xmin=12,xmax=Inf,ymin=0,ymax=Inf), alpha=0.1, fill="lightgrey", col="white") +
  geom_histogram(breaks=0:24, col="white", size=0.2) + coord_polar() +
  scale_x_continuous(breaks=seq(0,24,2)) +
  xlab("osc-modC acrophase")

```

```{r fig.height = 2,fig.width =3}
dat %>%
  # mutate(ageDMR = DMR_ENR_pval < 0.05) %>%
  filter(MHR_ENR_xCOLL_pval < 0.05) %>%
  ggplot(aes(x=MHR_ENR_xCOLL_acro)) +
  geom_rect(aes(xmin=12,xmax=Inf,ymin=0,ymax=Inf), alpha=0.1, fill="lightgrey", col="white") +
  geom_histogram(breaks=0:24, col="white", size=0.2) +
  scale_x_continuous(breaks=seq(0,24,4)) +
  xlab("osc-modC acrophase")

```

#### Osc-modC acrophase with IV

```{r fig.height = 2,fig.width =3}
dat %>%
  # mutate(ageDMR = DMR_ENR_pval < 0.05) %>%
  filter(MHR_ENRIV_xCOLL_pval < 0.05) %>%
  ggplot(aes(x=MHR_ENRIV_xCOLL_acro)) +
  geom_rect(aes(xmin=12,xmax=Inf,ymin=0,ymax=Inf), alpha=0.1, fill="lightgrey", col="white") +
  geom_histogram(breaks=0:24, col="white", size=0.2) +
  scale_x_continuous(breaks=seq(0,24,4)) +
  xlab("osc-modC acrophase")

```

#### Osc-modC histogram
```{r fig.height = 2,fig.width =3}
dat %>%
  ggplot(aes(x=MHR_ENR_xCOLL_pval)) +
  geom_histogram(breaks=0:20/20, col="white", size=0.2) +
  xlab("harmonic regression p-value")

```

#### Osc-modC histogram with IV
```{r fig.height = 2,fig.width =3}

dat %>%
  ggplot(aes(x=MHR_ENRIV_xCOLL_pval)) +
  geom_histogram(breaks=0:20/20, col="white", size=0.2) +
  xlab("harmonic regression p-value")


```

#### Pseudo-aging histogram
```{r fig.height = 2,fig.width =3}
dat %>%
  ggplot(aes(x=MPseudo2_ENR_pval)) +
  geom_histogram(breaks=0:20/20, col="white", size=0.2) +
  xlab("donor age linear regression p-value")

```

#### Ex vivo histogram
```{r fig.height = 2,fig.width =3}
dat %>%
  ggplot(aes(x=EXvivo_ENR_xCOLL_pval)) +
  geom_histogram(breaks=0:20/20, col="white", size=0.2) +
  xlab("epigenetic drift p-value")

```

#### Ex vivo drift
```{r fig.height = 2,fig.width =6}

dat %>%
  select(modid, EXvivo_ENR_16_pval, EXvivo_ENR_31_pval, EXvivo_ENR_62_pval) %>%
  melt(., id.var="modid") %>%
  mutate(variable = case_when(grepl("16", variable) ~"postnatal day 15",
                              grepl("62", variable) ~"postnatal day 60",
                           grepl("31", variable) ~"postnatal day 30")) %>%
  ggplot(aes(x=abs(value))) +  
  geom_histogram(breaks=0:20/20, size=0.2, col="white") +
  xlab("epigenetic drift p-value") +
  facet_wrap(~variable)

```

#### IV DMR
```{r fig.height = 2,fig.width =6}

dat %>%
  select(modid, IVDMR_16_pval, IVDMR_31_pval, IVDMR_62_pval) %>%
  melt(., id.var="modid") %>%
  mutate(variable = case_when(grepl("16", variable) ~"postnatal day 15",
                              grepl("62", variable) ~"postnatal day 60",
                           grepl("31", variable) ~"postnatal day 30")) %>%
  ggplot(aes(x=abs(value))) +  
  geom_histogram(breaks=0:20/20, size=0.2, col="white") +
  xlab("IV treatment p-value") +
  facet_wrap(~variable)

```

#### Cell cycle
```{r fig.height = 2,fig.width =6}

dat %>%
  select(modid, MCC_G2M_ENR_xCOLL_pval, MCC_S_ENR_xCOLL_pval) %>%
  melt(., id.var="modid") %>%
  mutate(variable = case_when(grepl("G2M", variable) ~"G2/M phase",
                              grepl("S", variable) ~"Synthesis phase")) %>%
  ggplot(aes(x=abs(value))) +  
  geom_histogram(breaks=0:20/20, size=0.2, col="white") +
  xlab("cell cycle p-value") +
  facet_wrap(~variable)

```
